---
title: "Defeo_Final"
author: "Shelby Defeo"
date: "4/25/2022"
output: word_document
---
Question 1:
Let’s explore the possible implications of these revised standards. For Merced County, is there a 
significant difference in the number of days in a year that exceed EPA ozone standards as 
defined by the 2008 NAAQ versus the 2015 NAAQ? If so, what is the magnitude of the 
difference? Base your analysis on the following years: 2021 and 2011. Note, in this question, you 
are not comparing 2021 to 2011 – we will do that in the next question.  
Provide a 250-500 word summary or your analysis and interpretation of your results. Please cite 
all outside references (e.g., papers, government reports) used to support your interpretation.
```{r}
##Question 1

##Import data and arrange to be able to compare ozone values and standards
library(tidyverse)
data_2001 <- read.csv("Ozone2001.csv")
data_2001
ozone_2001 <- (data_2001$Daily.Max.8.hour.Ozone.Concentration)
ozone_2001

data_2011 <- read.csv("Ozone2011.csv")
data_2011
ozone_2011 <- (data_2011$Daily.Max.8.hour.Ozone.Concentration)
ozone_2011

data_2021 <- read.csv("Ozone2021.csv")
data_2021
ozone_2021 <- (data_2021$Daily.Max.8.hour.Ozone.Concentration)
ozone_2021

##Compare 2001 to 2008 and 2015 standards
#2008 standard of 0.075
over2008_01 <- length(ozone_2001[ozone_2001>0.075])
over2008_01

#2015 standard of 0.70
over2015_01 <- length(ozone_2001[ozone_2001>0.07])
over2015_01

#Compare to assess for significant difference
binom.test(61, 84, p = 0.5,
           alternative = c("two.sided"),
           conf.level = 0.95)


##Compare 2011 to 2008 and 2015 standards
#2008 standard of 0.075
over2008_11 <- length(ozone_2011[ozone_2011>0.075])
over2008_11

#2015 standard of 0.70
over2015_11 <- length(ozone_2011[ozone_2011>0.07])
over2015_11

#Compare to assess for significant difference
binom.test(19, 38, p = 0.5,
           alternative = c("two.sided"),
           conf.level = 0.95)


##Compare 2021 to 2008 and 2015 standards
#2008 standard of 0.075
over2008_21 <- length(ozone_2021[ozone_2021>0.075])
over2008_21

#2015 standard of 0.70
over2015_21 <- length(ozone_2021[ozone_2021>0.07])
over2015_21

#Compare to assess for significant difference
binom.test(7, 21, p = 0.5,
           alternative = c("two.sided"),
           conf.level = 0.95)
```




Question 2:  
Is there a significant difference in ozone concentration between years (2001, 2011, 2021)? If so, 
what is possibly driving that difference. Note, for this analysis, you should conduct three 
separate comparisons: 1) 2001 & 2011, 2) 2011 & 2021, 3) 2001 & 2021. 
Provide a 250-500 word summary or your analysis interpretation of your results. Please cite all 
outside references (e.g., papers, government reports) used to support your interpretation. 

```{r}
##Question 2

##Import data and arrange to be able to compare ozone values
library(tidyverse)
data_2001 <- read.csv("Ozone2001.csv")
data_2001
ozone_2001 <- (data_2001$Daily.Max.8.hour.Ozone.Concentration)
ozone_2001

data_2011 <- read.csv("Ozone2011.csv")
data_2011
ozone_2011 <- (data_2011$Daily.Max.8.hour.Ozone.Concentration)
ozone_2011

data_2021 <- read.csv("Ozone2021.csv")
data_2021
ozone_2021 <- (data_2021$Daily.Max.8.hour.Ozone.Concentration)
ozone_2021


##Assess if 3 assumptions for parametric tests are met
qqnorm(ozone_2001)
qqline(ozone_2001)

qqnorm(ozone_2011)
qqline(ozone_2011)

qqnorm(ozone_2021)
qqline(ozone_2021)
#The three graphs each have slight deviations from normal, but I feel that they are close enough for me to assume normality.

sd01 <- sd(ozone_2001)
sd01

sd11 <- sd(ozone_2011)
sd11

sd21 <- sd(ozone_2021)
sd21

var01_11 <- sd01/sd11
var01_11

var01_21 <- sd01/sd21
var01_21

var21_11 <- sd21/sd11
var21_11

#The variance ratios are all within 0.5 and 2 so we can assume that the variance is approximately homogeneous.
#Assume from website/data that all the data were collected in a way that they were randomly and independently sampled.
#I will use a parameteric test because all 3 assumptions are met.

##Assess differences between ozone concentrations in years
t.test(ozone_2001, ozone_2011, mu = 0, paired = FALSE, var.equal = TRUE, conf.level = 0.95)
#p-value = 0.0005137

t.test(ozone_2011, ozone_2021, mu = 0, paired = FALSE, var.equal = TRUE, conf.level = 0.95)
#p-value = 0.005297

t.test(ozone_2001, ozone_2021, mu = 0, paired = FALSE, var.equal = TRUE, conf.level = 0.95)
#p-value = 5.929e-11

mean01 <- mean(ozone_2001)
mean01

mean11 <- mean(ozone_2011)
mean11

mean21 <- mean(ozone_2021)
mean21
```

